\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-10-12 doc/USER_MANUAL/03_mesh_generation.tex

\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
%
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
%
\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
that a $\nexxi=288$ run is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation will be accurate to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
\end{equation}
 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest not to change this value.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
is set to \texttt{.true.}.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}).
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).

>>>>

\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
%
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
%
\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
that a $\nexxi=288$ run is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation will be accurate to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
\end{equation}
 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest not to change this value.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}.
Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
For higher numbers of doubling layers, additional entries must be added.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
The layer count starts from the bottom layer.
More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).

### 2015-10-12 doc/USER_MANUAL/08_doing_tomography.tex
updates user manual

UNDER CONSTRUCTION (April 2015). Will be added around June 2015.

@@

two parameters. With this new parametrisation, the perturbation of the misfit function may be written as:

@@

Computing the gradient based on the adjoint-state method requires performing two simulations per source (forward and adjoint
fields) regardless of the type of observable. However, to define the adjoint field one
must know the adjoint source, and that source is computed from the results of the forward simulation. One must therefore perform
the forward simulation before the adjoint simulation.
A straightforward solution for time-domain methods would be to store the whole forward field to disk at
each time step during the forward run and then read it back during the adjoint simulation to calculate the interaction of these two fields.
In 2-D this is feasible but in the 3-D case for very short seismic periods and without lossy compression, downsampling,
or a large amount of disk or memory checkpointing \cite[e.g.,][]{FiKeIg09,RuHaPuGu14,CyShWi15}
the amount of disk storage required would currently be too large.
However let us note again that this situation will change in the future.
In the mean time, a standard possible solution is to perform three simulations per source \cite[]{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
i.e., perform the forward calculation twice, once to compute the adjoint sources and once again
at the same time as the adjoint simulation to correlate the two fields and sum their interaction on the fly over all the time steps.
Doing so for an elastic Earth, one only needs a small amount of disk storage to store the last time
step of the forward run, which is then used as an initial condition to redo the forward run backwards,
as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.

@@

There are dozens of versions of gradient-based inversion
algorithms that could alternatively be used. The tomographic inversion
of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D Cartesian as well
as several additional components which are also stored on the CIG
svn server, described next.

>>>>

UNDER CONSTRUCTION (July 2015). New content will be added soon.

@@

two parameters. With this new parametrization, the perturbation of the misfit function may be written as:

@@

Computing the gradient based on the adjoint-state method requires performing two simulations per source (forward and adjoint
fields) regardless of the type of observable. However, to define the adjoint field one
must know the adjoint source, and that source is computed from the results of the forward simulation. One must therefore perform
the forward simulation before the adjoint simulation.
A straightforward solution for time-domain methods would be to store the whole forward field to disk at
each time step during the forward run and then read it back during the adjoint simulation to calculate the interaction of these two fields.
In 2-D this is feasible but in the 3-D case for very short seismic periods and without lossy compression, downsampling,
or a large amount of disk or memory checkpointing \cite[e.g.,][]{FiKeIg09,RuHaPuGu14,CyShWi15}
the amount of disk storage required would currently be too large.

However let us note again that this situation will change in the future.
In the mean time, a standard possible solution is to perform three simulations per source \cite[]{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
i.e., perform the forward calculation twice, once to compute the adjoint sources and once again
at the same time as the adjoint simulation to correlate the two fields and sum their interaction on the fly over all the time steps.
Doing so for an elastic Earth, one only needs a small amount of disk storage to store the last time
step of the forward run, which is then used as an initial condition to redo the forward run backwards,
as well as the field on the outer edges of the mesh for each time steps in order to be able to undo the absorbing boundary conditions.

\section{Tomographic tools}

Besides the ability to compute adjoint sensitivity kernels, the SPECFEM3D package provides a number of tools useful for post-processing kernels, gradients and conducting tomographic model updates. The tomographic and post-processing tools can be compiled by typing in the main directory:
\begin{verbatim}
make tomography
make postprocess
\end{verbatim}
You can modify specific settings affecting, e.g., the (isotropic) parameterization or directory setup structure in file \texttt{setup/constants\_tomography.h}. The compiled binaries can be found in directory \texttt{bin/}.\\

\noindent
The iterative inversion workflow may comprise three main steps for which tools are provided:
\begin{enumerate}
\item Summing event kernels, e.g. $K_{\alpha,\beta,\rho}$, to build the misfit gradient $\boldg(\boldm)$
\item Smoothing and post-processing of the gradient $\boldg(\boldm)$
\item Updating the model $\boldm_{i+1} = \boldm_i + \delta\boldm$.
\end{enumerate}

Please note that the tomographic routines are very similar for both SPECFEM3D\_Cartesian and SPECFEM3D\_GLOBE versions. Some differences occur, e.g., when reading in mesh files and when the maximum of the gradient is taken for the update step length. Also note that currently only the binary database format is supported, no ADIOS file format support is implemented yet. In future, we plan to extend and merge these into the same set of tools within a common SPECFEM function library.\\

\subsection{Summing kernels}

For summing different event kernels, the executables \texttt{xsum\_kernels} and \texttt{xsum\_preconditioned\_kernels} can be used.
The binaries sum up either a set of isotropic or transversely isotropic kernel files, depending on the setting in file \texttt{setup/constants\_tomography.h}.
The following parameterizations can be chosen:
\begin{enumerate}
\item [-] isotropic kernels $K_{\alpha,\beta,\rho}$ or
\item [-] isotropic bulk kernels $K_{c,\beta,\rho}$, or
\item [-] transversely isotropic bulk kernels $K_{c,\beta_v,\beta_h,\eta}$.
\end{enumerate}

\noindent
The tools for summing event kernels can be used in parallel, using the same number of processes (\texttt{NPROC}) as the forward/kernel simulations.
The summation tools use the following input/output format (see Fig.~\ref{fig:tomo-dir-struct} for the default directory structure):
\begin{itemize}
\item [-]{\bf Input:}
Input is provided by a list of event directories, given in file \texttt{kernels\_list.txt} (default setting for \texttt{KERNEL\_FILE\_LIST} in \texttt{constants\_tomography.h}). The file lists all event kernel directories which should be summed together. All event directories have to be located in directory \texttt{INPUT\_KERNELS/}, which can be setup as links to all individual event kernel directories.

\item [-]{\bf Output:}
The summed kernels will be stored in corresponding kernel files in output directory \texttt{OUTPUT\_SUM/}.\\
\end{itemize}

\begin{figure}[htp]
\begin{centering}
\includegraphics[width=4in]{figures/tomo_dir_struct.pdf}
\par\end{centering}
%
\caption{
Example directory structure when using tomographic tools for summation. Event directory names (red) can be chosen freely with corresponding name entries in file \texttt{kernels\_list.txt}}
\label{fig:tomo-dir-struct}
\end{figure}

\begin{description}
\item [Kernel summation:] As example with a total of 4 MPI processes
\begin{verbatim}
mpirun -np 4 ./bin/xsum_kernels
\end{verbatim}
adds sensitivity kernels $K$ from different events together, i.e. outputs $\boldg$ where
\begin{equation}
\boldg = \Sum^{N}_i K^{i} \nonumber
\end{equation}
for events $i = 1,..,N$ as specified in file \texttt{kernel\_list.txt}.

\item [Kernel Summation with Preconditioning:] As example with a total of 4 MPI processes
\begin{verbatim}
mpirun -np 4 ./bin/xsum_preconditioned_kernels
\end{verbatim}
adds sensitivity kernels $K$ together from different events and "preconditions" the sum by dividing with the sum of the corresponding approximate Hessians $\tilde{H}$, i.e. outputs $\boldg$ where
\begin{equation}
\boldg = \frac{1}{\Sum_i \tilde{H}^{i}} \; \Sum_i  K^{i} \nonumber
\end{equation}
In this case, the factor $\frac{1}{\Sum_i \tilde{H}^{i}}$ acts as preconditioner, approximating the inverse of the Hessian $\mathbf{H}^{-1}$.
\end{description}

\noindent
{\bf Kernel names}: The summation tool will look for kernels with following names:
\begin{itemize}
\item [-] for an {\it isotropic} model parameterization, kernel names are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_alpha_kernel.bin
proc***_beta_kernel.bin
proc***_rho_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_alpha_kernel.bin
proc***_reg1_beta_kernel.bin
proc***_reg1_rho_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

\item [-] for an {\it isotropic bulk} model parameterization, kernel names are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_bulk_kernel.bin
proc***_bulk_beta_kernel.bin
proc***_rho_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_bulk_kernel.bin
proc***_reg1_bulk_beta_kernel.bin
proc***_reg1_rho_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

\item [-] for a {\it transversely isotropic} model parameterization, kernel names are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_bulk_c_kernel.bin
proc***_bulk_betav_kernel.bin
proc***_bulk_betah_kernel.bin
proc***_eta_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_bulk_c_kernel.bin
proc***_reg1_bulk_betav_kernel.bin
proc***_reg1_bulk_betah_kernel.bin
proc***_reg1_eta_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

Note that these event kernels are stored after a kernel simulation (\texttt{SIMULATION\_TYPE = 3}) in the \texttt{LOCAL\_PATH} directory, which by default points to directory \texttt{DATABASES\_MPI/}. Isotropic kernels will be created by default. To create transversely isotropic kernels, you set \texttt{ANISOTROPIC\_KL} and \texttt{SAVE\_TRANSVERSE\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}. \\

\item [- ] for preconditioning, the approximate Hessian kernels taken for summation are:\\
\begin{tabular}{ c c }
\begin{minipage}{3in}
\begin{verbatim}
## Cartesian version
proc***_hess_kernel.bin
\end{verbatim}
\end{minipage}
&
\begin{minipage}{3in}
\begin{verbatim}
## GLOBE version
proc***_reg1_hess_kernel.bin
\end{verbatim}
\end{minipage}
\end{tabular}\\

These Hessian kernels approximate the diagonal elements of the Hessian matrix and can be used as preconditioner in a gradient optimization scheme.
To create these kernels, you have to set \texttt{APPROXIMATE\_HESS\_KL} to \texttt{.true.} in the parameter file \texttt{DATA/Par\_file}.\\

\end{itemize}

For SPECFEM3D\_GLOBE, by default only kernels in the crust/mantle region ("reg1") will be considered for summation. You can change the default by setting parameter \texttt{REG} in file \texttt{constants\_tomography.h} to the region of interest. \\

Note that although we provide the preconditioned summation here, we recommend to smooth first both, the summed kernels and summed approximate Hessians, before inverting the summed Hessian and applying it as preconditioner to the gradient. From experience, we prefer doing
\begin{equation}
\boldg =  \frac{1}{\mathcal{F}_{smooth} \left( \Sum^{N}_i \tilde{H}^{i} \right) }  \; \mathcal{F}_{smooth} \left( \Sum_i K^{i}  \right) \nonumber
\end{equation}
rather than
\begin{equation}
\boldg = \mathcal{F}_{smooth} \left( \frac{1}{\Sum^{N}_i \tilde{H}^{i}}  \;  \Sum_i K^{i}  \right) \: . \nonumber
\end{equation}
Due to the large sensitivity kernel values close to source and receiver locations, the inverse of the thresholded summed Hessian becomes better balanced in the former.

\subsection{Smoothing and post-processing}
For additional smoothing and post-processing of kernels, gradients, or models, we provide tools useful for gradient and model preparation.
These additional tools for post-processing sensitivity kernels are provided together with the main package. For compilation of these tools, you type in the main directory:
\begin{verbatim}
make postprocess
\end{verbatim}

\noindent
The following tools are provided:
\begin{enumerate}
\item [-] \texttt{xsmooth\_sem} is used for smoothing with a gaussian function.
\item [-] \texttt{xclip\_sem} is used to threshold a kernel, gradient or model, clipping to a minimum and maximum value.
\item [-] \texttt{xcombine\_sem} is used to combine different event kernels, gradient or model files, summing all individually specified files together (similar to \texttt{xsum\_kernels} but more flexible).
\end{enumerate}
The post-processing tools are primarily intented to be used to process kernel files. They can be used though on
any scalar field of dimension \texttt{(NGLLX,NGLLY,NGLLZ,NSPEC)}.
The tools are parallel programs -- they must be invoked with mpirun or other
appropriate utility.  Operations are performed in embarrassingly-parallel fashion.

\begin{description}
\item [Smoothing:] To smooth a kernel, gradient or model file, use
\begin{verbatim}
mpirun -np NPROC bin/xsmooth_sem SIGMA_H SIGMA_V KERNEL_NAME INPUT_DIR OUPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{SIGMA\_H} horizontal smoothing radius,
\texttt{SIGMA\_V} vertical smoothing radius,
\texttt{KERNEL\_NAME} kernel name, e.g. \texttt{alpha\_kernel},
\texttt{INPUT\_DIR} directory from which kernels are read,
\texttt{OUTPUT\_DIR} directory to which smoothed kernels are written.

Smooths kernels by convolution with a Gaussian. Writes the resulting
smoothed kernels to \texttt{OUTPUT\_DIR}.

Files written to \texttt{OUTPUT\_DIR} have the suffix 'smooth' appended,
e.g. \texttt{proc***alpha\_kernel.bin} becomes \texttt{proc***alpha\_kernel\_smooth.bin}.

\item [Clipping:] Values in a kernel, gradient or model binary file can be clipped with a minimum/maximum threshold value using
\begin{verbatim}
mpirun -np NPROC bin/xclip_sem MIN_VAL MAX_VAL KERNEL_NAMES INPUT_FILE OUTPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{MIN\_VAL} threshold below which array values are clipped,
\texttt{MAX\_VAL} threshold above which array values are clipped,
\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
\texttt{INPUT\_DIR} directory from which arrays are read,
\texttt{OUTPUT\_DIR} directory to which clipped array are written.

For each name in \texttt{KERNEL\_NAMES}, reads kernels from \texttt{INPUT\_DIR}, applies
thresholds, and writes the resulting clipped kernels to \texttt{OUTPUT\_DIR}.

\texttt{KERNEL\_NAMES} is a comma-delimited list of kernel names,
e.g. \texttt{alphav\_kernel,alphah\_kernel}.
\begin{itemize}
\item [] For example, in this case type
{\small
\begin{verbatim}
mpirun -np 4 ./bin/xclip_sem -0.1 0.1 alphav_kernel,alphah_kernel DIR1/ DIR1/
\end{verbatim}
}
to clip the corresponding kernels in directory \texttt{DIR1} to be in range $[-0.1,0.1]$.
\end{itemize}

Files written to \texttt{OUTPUT\_DIR} have the suffix 'clip' appended,
e.g. \texttt{proc***alpha\_kernel.bin} becomes \texttt{proc***alpha\_kernel\_clip.bin}

\item [Combining:] To combine kernel values from different kernel directories, you can use
\begin{verbatim}
mpirun -np NPROC bin/xcombine_sem KERNEL_NAMES INPUT_FILE OUTPUT_DIR
\end{verbatim}
with command line arguments:
\texttt{KERNEL\_NAMES} one or more kernel names separated by commas,
\texttt{INPUT\_FILE} text file containing list of kernel directories,
\texttt{OUTPUT\_PATH} directory to which summed kernels are written.

For each name in \texttt{KERNEL\_NAMES}, sums kernels from directories specified in
\texttt{INPUT\_FILE}. Writes the resulting sums to \texttt{OUTPUT\_DIR}.

\texttt{INPUT\_FILE} is a text file containing a list of absolute or relative paths to
kernel directories, one directory per line.

\texttt{KERNEL\_NAMES} is a comma-delimited list of kernel names,
e.g. \texttt{alpha\_kernel,beta\_kernel,rho\_kernel}.

\end{description}

\subsection{Model updating}
We provide (simple) tools for updating the current model $\boldm_i$ using a gradient $\boldg$ and steplength $\alpha$:, i.e.
\begin{equation}
\boldm_{i+i} = \boldm_i + \delta \boldm  \nonumber
\end{equation}
with
\begin{equation}
\delta \boldm = - \alpha \, \boldg \nonumber \: .
\end{equation}
\noindent
The following tomographic tools are provided:
\begin{enumerate}
\item [-] \texttt{xadd\_model\_iso} can be used to update isotropic model files with a (summed \& smoothed) gradient.
The gradient files are given for isotropic parameters or isotropic bulk parameters.

\end{enumerate}

Note that instead of using the density kernels $K_{\rho}$ for model updates, setting \texttt{USE\_RHO\_SCALING} to \texttt{.true.} will
ignore the density kernel and use density perturbations scaled from isotropic $V_s$ pertubations.\\

\begin{description}
\item [Isotropic model update:] The program \texttt{xadd\_model\_iso} can be used to update {\it isotropic} model files with
(smoothed \& summed) event kernels:
\begin{verbatim}
mpirun -np 4 bin/xadd_model_iso step_factor [INPUT-KERNELS-DIR/] [OUTPUT-MODEL-DIR/]
\end{verbatim}
with command line arguments:
\texttt{step\_factor} the step length to scale the gradient, e.g. $0.03$ for a $\pm 3$\% update,
\texttt{INPUT-KERNELS-DIR/} (optional) directory which holds summed kernels (e.g. \texttt{proc***alpha\_kernel.bin},.., by default
directory \texttt{INPUT\_GRADIENT/} is used),
\texttt{OUTPUT-MODEL-DIR/} (optional) directory which will hold new model files (e.g. \texttt{proc***vp\_new.bin},.., by default directory
\texttt{OUTPUT\_MODEL/} is used).

The gradients are given for isotropic parameters \texttt{(alpha,beta,rho)} or \texttt{(bulk\_c,beta,rho)}.

The algorithm uses a steepest descent method with a step length
determined by the given maximum update percentage.

By default, the directory and file setup used is:
\begin{itemize}
\item [-] directory \texttt{INPUT\_MODEL/} contains:
\begin{verbatim}
proc***_vs.bin
proc***_vp.bin
proc***_rho.bin
\end{verbatim}

\item [-] directory \texttt{INPUT\_GRADIENT/} contains:
\begin{verbatim}
proc***_bulk_c_kernel_smooth.bin
proc***_bulk_beta_kernel_smooth.bin
proc***_rho_kernel_smooth.bin
\end{verbatim}
or
\begin{verbatim}
proc***_alpha_kernel_smooth.bin
proc***_beta_kernel_smooth.bin
proc***_rho_kernel_smooth.bin
\end{verbatim}
depending on the model parameterization.

\item [-] directory \texttt{topo/} contains:
\begin{verbatim}
proc***_external_mesh.bin
\end{verbatim}
for the Cartesian version, and
\begin{verbatim}
proc***_solver_data.bin
\end{verbatim}
for the GLOBE version.
\end{itemize}

\noindent
The new model files are stored in directory \texttt{OUTPUT\_MODEL/} as
\begin{verbatim}
proc***_vp_new.bin
proc***_vs_new.bin
proc***_rho_new.bin
\end{verbatim}

\end{description}

Note that additional post-processing and model update tools are provided for the SPECFEM3D\_GLOBE version and will be provided for the Cartesian version as well in future.

@@

There are dozens of versions of gradient-based inversion
algorithms that could alternatively be used. The tomographic inversion
of \citet{TaLiMaTr09,TaLiMaTr2010} used SPECFEM3D\_Cartesian as well
as several additional components which are also stored on the CIG
svn server, described next.
